---
title: "Glucose Diary of April 2017"
output: 
  html_document
---

```{r init, echo = F, message=F, warning=F}
## knitr options setzen
knitr::opts_chunk$set(echo = F, warning = F, message = F, fig.align = "center",
                      fig.path = "assets/apr2017_files/")
options(digits = 3)


## toolbox etc
library(tidyverse)
library(sjPlot)
library(plotly)
library(ggthemes)
library(viridis)
library(tadaatoolbox)


## read data
sugr    <- readRDS("./data/all_sugr.rds") %>% filter(Monat == "April", Jahr == "2017")
base    <- readRDS("./data/base.rds")
bexpert <- readRDS("./data/all_bexpert.rds") %>% filter(Monat == "April", Jahr == "2017")


## prepare cosmetics
source("./assets/theme_tut.R")
theme_set(theme_tut(size = 14))
```

# Overviews

## Averages

```{r overview_table}
by_days <-  sugr %>% 
  group_by(Datum) %>% 
  summarize(
    mean_bz = mean(BZ_Wert, na.rm = T),
    sum_be  = sum(BE_KH, na.rm = T),
    sum_ie  = sum(BE_Schaetzung, na.rm = T),
    sum_kor = sum(BE_Korrektur, na.rm = T)
    )

by_days %>% 
  summarize(
    "∅ Glucose Level (mg/dl)" = mean(mean_bz), 
    "∅ BE" = mean(sum_be), 
    "∅ IU" = mean(sum_ie),
    "% of IU for Correction" = mean(sum_kor) / mean(sum_ie) * 100
    ) %>% 
  knitr::kable(caption = "Daily averages")
```

## Monthly development

```{r table_month}
sugr %>% 
  filter(!is.na(BZ_Wert)) %>% 
  summarise(
    "∅ Glucose Level (mg/dl)" = mean(BZ_Wert),
    "SD of Glucose Level"     = sd(BZ_Wert),
    "Median of Glucose Level" = median(BZ_Wert),
    "Measurements"            = length(BZ_Wert)
  ) %>% 
  knitr::kable(digits = 0, caption = "Monthly Stats")
```

```{r overview_month}
outliers <- sugr %>% 
  filter(!is.na(BZ_Wert), BZ_Wert > 350) %>% 
  summarize(n = n()) %>% 
  .$n

ggplot(sugr[!is.na(sugr$BZ_Wert) & sugr$BZ_Wert < 350,], 
       aes(x = Datum, y = BZ_Wert)) + 
  geom_hline(yintercept = c(75, 175), color = "red", size = .5, linetype = "dashed") +
  geom_jitter(width = .2, size = 1.5, alpha = .7, color = "#9C3848",
              stroke = .8, fill = "white", shape = 21) +
  geom_smooth(method = "loess", color = "#1E3888", fill = "#1E3888", alpha = 0.3) +
  scale_x_date(date_breaks = "1 week", date_minor_breaks = "1 day",
               date_labels = "%b '%d") +
  labs(title = paste0("Overview of ", unique(sugr$Monat), " ", unique(sugr$Jahr), "¹"),
       subtitle = "red dashes show target range",
       y = "glucose level (mg/dl)", x = "Date",
       caption = paste("¹", outliers, "measurements above 350 mg/dl taken out"))
```
I count 6 days with no measurements below 80 mg/dl. Quite depressing. Way too many measurements on the upper end, too, albeit those can be explained by the usual hiccups: counter-reactions to low levels, miscalculated meals, alcohol, and so on.

## Glucose ranges by day

```{r table_days}
sugr %>% 
  filter(!is.na(BZ_Wert)) %>% 
  group_by(Wochentag) %>% 
  summarise(
    "∅ Glucose Level (mg/dl)" = mean(BZ_Wert),
    "SD of Glucose Level"     = sd(BZ_Wert),
    "Median of Glucose Level" = median(BZ_Wert),
    # "Measurements below target range" = 
    "Measurements"            = length(BZ_Wert)
  ) %>% 
  knitr::kable(digits = 0, caption = "Daily Stats")
```

```{r overview_boxplot}
ggplot(sugr[!is.na(sugr$BZ_Wert) & sugr$BZ_Wert < 350,], 
       aes(x = Wochentag, y = BZ_Wert, color = Wochentag, fill = Wochentag)) +
  geom_hline(yintercept = c(75, 175), color = "red", size = .5, linetype = "dashed") +
  geom_jitter(width = .2, alpha = .5, size = 1.5, stroke = 1, fill = "white", shape = 21) +
  geom_boxplot(alpha = .1) +
  scale_fill_viridis(discrete = T, option = "C") +
  scale_color_viridis(discrete = T, option = "C") +
  labs(title = "Overview of glucose ranges by day¹", y = "glucose level (mg/dl)",
       subtitle = "red dashes show target range", x = NULL,
       caption = paste("¹", outliers, "measurements above 350 mg/dl taken out")) +
  theme(legend.position = "none")
```
I wonder what makes tuesdays as stable as they are?

## Bolus-Base-Ratio by day

```{r bbratio}
sugr %>% 
  mutate(
    Bolus_Gesamt = BE_Schaetzung + BE_Korrektur
  ) %>% 
  group_by(Datum, Wochentag) %>% 
  summarise(
    p_bolus = sum(Bolus_Gesamt, na.rm = T) / (sum(Bolus_Gesamt, na.rm = T) + 12.05),
    p_basal = 1 - p_bolus
  ) %>% 
  ungroup() %>% 
  group_by(Wochentag) %>% 
  summarise(
    Bolus = mean(p_bolus),
    Basal = mean(p_basal)
  ) %>% 
  gather(Insulinanteil, Prozent, Bolus, Basal) %>% 
  ggplot(aes(x = Wochentag, y = Prozent, fill = Insulinanteil)) +
    geom_bar(stat = "identity", alpha = .6) +
    scale_y_continuous(labels = scales::percent, position = "right") +
    scale_fill_viridis(discrete = T, direction = -1) +
    labs(x = NULL, y = NULL, fill = NULL,
         title = "Proportions of Base to Bolus") +
    coord_flip()
```

# Closer look at measurements

## All measurements

Let's take a look at the development of an average day. Since I started my apprenticeship my night hours shifted somewhat from 0 to 7 o'clock. 

```{r BZ_Gesamt}
sugr %>% 
  filter(BZ_Wert < 350, !is.na(BZ_Wert)) %>% 
  ggplot(aes(x = Zeit, y = BZ_Wert)) + 
    annotate("rect", fill = "black", alpha = .1,
             xmin = 1, xmax = 25200, ymin = 0, ymax = Inf) +
    geom_hline(yintercept = c(75, 175), color = "red", size = .5, linetype = "dashed") +
    geom_point(size = 1.5, alpha = .7, color = "#9C3848",
               stroke = .8, fill = "white", shape = 21) +
    geom_smooth(method = "loess", color = "#1E3888", fill = "#1E3888", alpha = 0.3) +
    scale_x_continuous(labels = paste0(seq(0, 24, 4), ":00 h"),
                       breaks = seq(0, 86400, 14400), 
                       minor_breaks = seq(0, 86400, 7200)) +
    labs(title = "Average daily glucose development¹", x = "Time", y = "glucose level (mg/dl)",
         caption = paste("¹", outliers, "measurements above 350 mg/dl taken out"))
```

Now this is seriously unsettling. I'm mostly fine while awake it seems (well, apart from all the hypos...), but before going to bed and when getting up again, my levels are way too high!

## by weekday

Let's continue with looking at the weekdays. It's only reasonable to expect different developments on different days (i.e. working days and weekends).

```{r BZ_pro_Tag, fig.height=12}
sugr %>% 
  filter(BZ_Wert < 350, !is.na(BZ_Wert)) %>% 
  ggplot(aes(x = Zeit, y = BZ_Wert, color = BZ_Wert)) + 
    annotate("rect", fill = "black", alpha = .1,
             xmin = 1, xmax = 25200, ymin = 0, ymax= Inf) +
    geom_point(size = 1.5, alpha = .7, color = "#9C3848",
               stroke = 1, fill = "white", shape = 21) +
    geom_hline(yintercept = c(80, 160), color = "red", size = .5, linetype = "dashed") +
    geom_smooth(method = "loess", color = "#1E3888", fill = "#1E3888", alpha = 0.3) +
    scale_x_continuous(labels = paste0(seq(0, 24, 4), ":00 h"),
                       breaks = seq(0, 86400, 14400), 
                       minor_breaks = seq(0, 86400, 7200)) +
    scale_y_continuous(limits = c(0, NA)) +
    labs(title = "Average glucose development by weekday¹", x = "Time", y = "glucose level (mg/dl)",
         caption = paste("¹", outliers, "measurements above 350 mg/dl taken out")) +
    facet_wrap(~Wochentag, ncol = 2)
```

Since I put aside high levels in the evening due to alcohol before, this is rather unexpected: the most typical day to have a drink (or two) to me is **Tuesday** (don't ask, student's life, yaddayadda...). Fri- & Saturdays too, but this month not so much because of the exam period. Also, come to think of it, when going out I tend to not look at my glucose level so much (or at all) since I already _know_ them to be high (or at least I do expect them to be).  
The biggest problem to me seems to be the overall tendency to end the day with too high levels, so without further trying to read the cards, I will tackle that problem first.

## Weird _BolusExpert_-measures

Now this is really somethring I'm scratching my head over: the pump seems to have two ways to measure glucose levels (or at least there are two columns with different timestamps and values). The first is a somewhat _"regular"_ column internally called ``r sjmisc::get_label(sugr$BZ_Wert)[[1]]`` and another one called ``r sjmisc::get_label(sugr$BE_BZ)[[1]]`` (translated _"GL-Measurement"_ and _"BolusExpert: GL-Input"_ respectively).  
This much I get: _"BolusExpert"_ is the name of the little helper software of the pump, which does calculation of needed IUs or correction-IUs and factors in already/still active insulin and such. Sometimes I just take a measurement and don't go on with anything else; sometimes I just want to eat something but don't look at my GL first - this at least would explain _"regular"_ measurements without _"Bolusexpert"_-entries and _BE_-GL-values of zero.  

What really bothers me, though, is the difference between the two measurements:

```{r BZ_BE_tbl}
t <- sugr %>% 
  gather(Measurement, Wert, BZ_Wert, BE_BZ) %>% 
  mutate(Measurement = recode(Measurement, "BE_BZ" = "BolusExpert", "BZ_Wert" = "Regular")) %>% 
  filter(Wert > 0)

t %>% 
  group_by(Measurement) %>% 
  summarize(N    = n(),
            Mean = mean(Wert),
            SEM  = sd(Wert)/sqrt(n()),
            SD   = sd(Wert)) %>% 
  knitr::kable(digits = 2)
```

```{r BZ_BE_timeline}
ggplot(t, aes(x = Zeit, y = Wert, color = Measurement, fill = Measurement)) +
  annotate("rect", fill = "black", alpha = .1,
           xmin = 1, xmax = 25200, ymin = 0, ymax= Inf) +
  geom_hline(yintercept = c(80, 160), color = "red", size = .5, linetype = "dashed") +
  geom_point(size = 1.5, alpha = .5, stroke = 1, fill = "white", shape = 21) +
  geom_smooth(method = "loess", alpha = .2) +
  scale_x_continuous(labels = paste0(seq(0, 24, 4), ":00 h"),
                     breaks = seq(0, 86400, 14400), 
                     minor_breaks = seq(0, 86400, 7200)) +
  labs(x = "Time", y = "glucose level (mg/dl)") +
  scale_color_viridis(discrete = T) +
  scale_fill_viridis(discrete = T)
```

```{r BZ_BE_bar}
t %>% 
  group_by(Measurement) %>% 
  summarize(N    = n(),
            Mean = mean(Wert),
            SEM  = sd(Wert)/sqrt(n()),
            SD   = sd(Wert)) %>% 
ggplot(aes(x = Measurement, y = Mean, color = Measurement, fill = Measurement)) +
  geom_bar(stat = "identity", alpha = 0.4) +
  geom_errorbar(aes(ymin = SD * -1 + Mean, ymax = SD + Mean),
                size = 1, width = 0.7) +
  labs(y = "glucose level (mg/dl)",
       title = "Mean of Glucoselevel by Measurement-Method",
       subtitle = "Errorbars depicting SD") +
  scale_color_viridis(discrete = T) +
  scale_fill_viridis(discrete = T)
```

While their SDs and distribution stay roughly the same, the mean of both methods is almost 90 mg/dl apart. This is huge! While the difference in number of uses is most probably explained, as I said earlier, by measured levels that don't need any further attention (and thus no use of the _BolusExpert_), my only explanation to the huge difference in glucose levels is just the opposite of that: If mostly high levels get piped through _BE_ (i.e. for a corrective bolus output), it seems reasonable for the mean being so much higher.

What is still bothering is the following:  
Most _regular_ measurement entries get followed by an comparable or equal entry in the _BE_ column shortly after. This is to be expectet, since I need maybe a minute to properly react (i.e. calculate carbohydrates or just handling the pump). Curiously though, **_not every_** high _BE_-entry is preceded by an equally high _regular_ entry (comment on 04.03.17: _actually, I have to look into that; maybe the problem just resolves itself if it is, after all_). Why is that? What? How? Sad!
