---
title: "Glucose Diary of March 2017"
output: 
  html_document
---

```{r init, echo = F, message=F, warning=F}
## knitr options setzen
knitr::opts_chunk$set(echo = F, warning = F, message = F, fig.align = "center",
                      fig.path = "assets/mar2017_files/")
options(digits = 3)


## toolbox etc
library(tidyverse)
library(sjPlot)
library(plotly)
library(ggthemes)
library(viridis)
library(tadaatoolbox)


## read data
sugr    <- readRDS("./data/all_sugr.rds") %>% filter(Monat == "March", Jahr == "2017")
base    <- readRDS("./data/base.rds")
bexpert <- readRDS("./data/all_bexpert.rds") %>% filter(Monat == "March", Jahr == "2017")


## prepare cosmetics
source("./assets/theme_tut.R")
theme_set(theme_tut(size = 12))
```

# Overviews

## Averages

```{r overview_table}
by_days <-  sugr %>% 
  group_by(Datum) %>% 
  summarize(
    mean_bz = mean(BZ_Wert, na.rm = T),
    sum_be  = sum(BE_KH, na.rm = T),
    sum_ie  = sum(BE_Schaetzung, na.rm = T),
    sum_kor = sum(BE_Korrektur, na.rm = T)
    )

by_days %>% 
  summarize(
    "∅ Glucose Level (mg/dl)" = mean(mean_bz), 
    "∅ BE" = mean(sum_be), 
    "∅ IU" = mean(sum_ie),
    "% of IU for Correction" = mean(sum_kor) / mean(sum_ie) * 100
    ) %>% 
  knitr::kable(caption = "Daily averages")
```

## Monthly development

```{r overview_month}
outliers <- sugr %>% 
  filter(!is.na(BZ_Wert), BZ_Wert > 350) %>% 
  summarize(n = n()) %>% 
  .$n

ggplot(sugr[!is.na(sugr$BZ_Wert) & sugr$BZ_Wert < 350,], 
       aes(x = Datum, y = BZ_Wert)) + 
  geom_hline(yintercept = c(75, 175), color = "red", size = .5, linetype = "dashed") +
  geom_jitter(width = .2, alpha = .5, size = 1.5, color = "#66cc00") +
  geom_smooth(method = "loess", se = T, color = "#9966ff", fill = "#9966ff", alpha = .2) +
  scale_x_date(date_breaks = "1 week", date_minor_breaks = "1 day",
               date_labels = "%b '%d") +
  labs(title = paste0("Overview of ", unique(sugr$Monat), " ", unique(sugr$Jahr), "¹"),
       subtitle = "red dashes show target range",
       y = "glucose level (mg/dl)", x = "Date",
       caption = paste("¹", outliers, "measurements above 350 mg/dl taken out"))
```

## Glucose ranges by day

```{r overview_boxplot}
ggplot(sugr[!is.na(sugr$BZ_Wert) & sugr$BZ_Wert < 350,], 
       aes(x = Wochentag, y = BZ_Wert, color = Wochentag, fill = Wochentag)) +
  geom_hline(yintercept = c(75, 175), color = "red", size = .5, linetype = "dashed") +
  geom_boxplot(alpha = .1) +
  geom_jitter(width = .2, alpha = .4, size = 1.5) +
  scale_fill_viridis(discrete = T, option = "C") +
  scale_color_viridis(discrete = T, option = "C") +
  labs(title = "Overview of glucose ranges by day¹", y = "glucose level (mg/dl)",
       subtitle = "red dashes show target range", x = NULL,
       caption = paste("¹", outliers, "measurements above 350 mg/dl taken out")) +
  theme(legend.position = "none")
```

# Closer look at measurements

## All measurements

Let's take a look at the development of an average day:

```{r BZ_Gesamt}
sugr %>% 
  filter(BZ_Wert < 350, !is.na(BZ_Wert)) %>% 
  ggplot(aes(x = Zeit, y = BZ_Wert)) + 
    annotate("rect", fill = "black", alpha = .1,
             xmin = 7201, xmax = 28800, ymin = 0, ymax = Inf) +
    geom_point(size = 1.5, alpha = .4, color = "#66cc00") +
    geom_smooth(method = "loess", color = "#9966ff", fill = "#9966ff", alpha = 0.2) +
    geom_hline(yintercept = c(75, 175), color = "red", size = .5, linetype = "dashed") +
    scale_x_continuous(labels = paste0(seq(0, 24, 4), ":00 h"),
                       breaks = seq(0, 86400, 14400), 
                       minor_breaks = seq(0, 86400, 7200)) +
    labs(title = "Average daily glucose development¹", x = "Time", y = "glucose level (mg/dl)",
         caption = paste("¹", outliers, "measurements above 350 mg/dl taken out"))
```

On average (the line and it's "shadow") it looks pretty okay, even though levels are a bit too high over all; especially in the evening. hHaving a closer look at the actual measurements additionally shows too many high levels when getting up in the morning (or what counts as such for a student) and an unsettling mix of too many too high as well as too low levels in the early afternoon.  
Most high levels in the evening should be explainable by alcohol consumption (I'll have a look at that in the next section), whereas the many lows later on happen independently. The next step should therefore be checking my basal rate configuration for that time period.

## by weekday

Let's continue with looking at the weekdays. It's only reasonable to expect different developments on different days (i.e. working days and weekends).

```{r BZ_pro_Tag, fig.height=12}
sugr %>% 
  filter(BZ_Wert < 350, !is.na(BZ_Wert)) %>% 
  ggplot(aes(x = Zeit, y = BZ_Wert, color = BZ_Wert)) + 
    annotate("rect", fill = "black", alpha = .1,
             xmin = 7201, xmax = 28800, ymin = 0, ymax= Inf) +
    geom_point(size = 1.5, alpha = .4, color = "#66cc00") +
    geom_smooth(method = "loess", color = "#9966ff", fill = "#9966ff", alpha = 0.2) +
    geom_hline(yintercept = c(80, 160), color = "red", size = .5, linetype = "dashed") +
    scale_x_continuous(labels = paste0(seq(0, 24, 4), ":00 h"),
                       breaks = seq(0, 86400, 14400), 
                       minor_breaks = seq(0, 86400, 7200)) +
    scale_y_continuous(limits = c(0, NA)) +
    labs(title = "Average glucose development by weekday¹", x = "Time", y = "glucose level (mg/dl)",
         caption = paste("¹", outliers, "measurements above 350 mg/dl taken out")) +
    facet_wrap(~Wochentag, ncol = 2)
```

Since I put aside high levels in the evening due to alcohol before, this is rather unexpected: the most typical day to have a drink (or two) to me is **Tuesday** (don't ask, student's life, yaddayadda...). Fri- & Saturdays too, but this month not so much because of the exam period. Also, come to think of it, when going out I tend to not look at my glucose level so much (or at all) since I already _know_ them to be high (or at least I do expect them to be).  
The biggest problem to me seems to be the overall tendency to end the day with too high levels, so without further trying to read the cards, I will tackle that problem first.

## Weird _BolusExpert_-measures

Now this is really somethring I'm scratching my head over: the pump seems to have two ways to measure glucose levels (or at least there are two columns with different timestamps and values). The first is a somewhat _"regular"_ column internally called ``r sjmisc::get_label(sugr$BZ_Wert)[[1]]`` and another one called ``r sjmisc::get_label(sugr$BE_BZ)[[1]]`` (translated _"GL-Measurement"_ and _"BolusExpert: GL-Input"_ respectively).  
This much I get: _"BolusExpert"_ is the name of the little helper software of the pump, which does calculation of needed IUs or correction-IUs and factors in already/still active insulin and such. Sometimes I just take a measurement and don't go on with anything else; sometimes I just want to eat something but don't look at my GL first - this at least would explain _"regular"_ measurements without _"Bolusexpert"_-entries and _BE_-GL-values of zero.  

What really bothers me, though, is the difference between the two measurements:

```{r BZ_BE_tbl}
t <- sugr %>% 
  gather(Measurement, Wert, BZ_Wert, BE_BZ) %>% 
  mutate(Measurement = recode(Measurement, "BE_BZ" = "BolusExpert", "BZ_Wert" = "Regular")) %>% 
  filter(Wert > 0)

t %>% 
  group_by(Measurement) %>% 
  summarize(N    = n(),
            Mean = mean(Wert),
            SEM  = sd(Wert)/sqrt(n()),
            SD   = sd(Wert)) %>% 
  knitr::kable(digits = 2)
```

```{r BZ_BE_timeline}
ggplot(t, aes(x = Zeit, y = Wert, color = Measurement, fill = Measurement)) +
  annotate("rect", fill = "black", alpha = .1,
           xmin = 7201, xmax = 28800, ymin = 0, ymax= Inf) +
  geom_smooth(method = "loess", alpha = .2) +
  geom_point(size = 1.5, alpha = .5) +
  geom_hline(yintercept = c(80, 160), color = "red", size = .5, linetype = "dashed") +
  scale_x_continuous(labels = paste0(seq(0, 24, 4), ":00 h"),
                     breaks = seq(0, 86400, 14400), 
                     minor_breaks = seq(0, 86400, 7200)) +
  labs(x = "Time", y = "glucose level (mg/dl)") +
  scale_color_viridis(discrete = T) +
  scale_fill_viridis(discrete = T)
```

```{r BZ_BE_bar}
t %>% 
  group_by(Measurement) %>% 
  summarize(N    = n(),
            Mean = mean(Wert),
            SEM  = sd(Wert)/sqrt(n()),
            SD   = sd(Wert)) %>% 
ggplot(aes(x = Measurement, y = Mean, color = Measurement, fill = Measurement)) +
  geom_bar(stat = "identity", alpha = 0.4) +
  geom_errorbar(aes(ymin = SEM * -1 + Mean, ymax = SEM + Mean),
                size = 1, width = 0.7) +
  labs(y = "glucose level (mg/dl)",
       title = "Mean of Glucoselevel by Measurement-Method",
       subtitle = "With Errorbars depicting the SEM") +
  scale_color_viridis(discrete = T) +
  scale_fill_viridis(discrete = T)
```

While their SDs and distribution stay roughly the same, the mean of both methods is almost 90 mg/dl apart. This is huge! While the difference in number of uses is most probably explained, as I said earlier, by measured levels that don't need any further attention (and thus no use of the _BolusExpert_), my only explanation to the huge difference in glucose levels is just the opposite of that: If mostly high levels get piped through _BE_ (i.e. for a corrective bolus output), it seems reasonable for the mean being so much higher.

What is still bothering is the following:  
Most _regular_ measurement entries get followed by an comparable or equal entry in the _BE_ column shortly after. This is to be expectet, since I need maybe a minute to properly react (i.e. calculate carbohydrates or just handling the pump). Curiously though, **_not every_** high _BE_-entry is preceded by an equally high _regular_ entry (comment on 04.03.17: _actually, I have to look into that; maybe the problem just resolves itself if it is, after all_). Why is that? What? How? Sad!



# _here be dragons:_ Bolus & BEs 

This is really just fiddling around with the data; most doesn't make too much sense or is actually helpful to my analysis (but rather to my understanding of how the pump stores data). Therefore I won't even bother to update or translate it (for now). I will leave it here (also, for now) though; maybe it'll be useful?

## Bolusarten

Unter _"Bolusarten"_ verstehe ich die drei verschiedenen Modi, in denen die Pumpe die Abgabe festhält:

1. die von _BolusExpert_ geschätzte Insulinmenge, die benötigt wird
2. die von der Pumpe angeforderte Menge Insulin
3. die **tatsächlich** abgegebene Menge Insulin

Die angeforderte und die tatsächlich abgegebene Menge ist erwatungsgemäß Deckungsgleich ($r =$ `r cor(sugr$Bolus_abgegeben, sugr$Bolus_gewaehlt, use = "complete.obs")`), dementsprechend werden für die weiteren Beobachtungen jeweils die Werte der tatsächlich abgegebenen Menge verwendet.  
Spannender ist der Unterschied zwischen der _BolusExpert_-Schätzung und der abgegebenen Menge: Zum einen sind die Werte erstaunlich oft **nicht** identisch und zum anderen tauchen mal diese und mal jene auf:

```{r Bolusarten_Messungen}
sugr %>% 
  gather(Messung, Bolus, Bolus_abgegeben, BE_Schaetzung) %>% 
  select(Zeitstempel, Messung, Bolus) %>% 
  filter(!is.na(Bolus)) %>% 
  mutate(Messung = recode(Messung, "BE_Schaetzung" = "BolusExpert", "Bolus_abgegeben" = "tatsächlich abgegeben")) %>% 
  ggplot(aes(x = Zeitstempel, y = Bolus, color = Messung)) + 
    geom_point(alpha = 0.5) +
    labs(title = "Abgegebene Menge Insulin",
         subtitle = "nach Art der Messung",
         x = "Datum", y = "Bolus (IE)") +
    scale_color_viridis(discrete = T)
```

Was genau der Unterschied zwischen den beiden Varianten ist, ist mir noch unklar. Ein t-Test auf die Modi ergibt zumindest weder ein signifikantes Ergebnis, noch eine nennenswerte Effektröße, und das trotz der verhältnismäßig großen Gruppen von jeweils etwa 235 Beobachtungen:

```{r Bolusarten_t_test}
t_test <- sugr %>% 
  gather(Messung, Bolus, Bolus_abgegeben, BE_Schaetzung) %>% 
  mutate(Messung = recode(Messung, "BE_Schaetzung" = "BolusExpert", "Bolus_abgegeben" = "tatsächlich abgegeben")) %>% 
  filter(!is.na(Bolus))

tadaa_t.test(t_test, Bolus, Messung, print = "markdown")
rm(t_test)
```


