---
title: "Welcome to my Glucose Diary"
output:
  html_document:
    toc: false
---

```{r init, echo = F, message=F, warning=F}
## knitr options setzen
knitr::opts_chunk$set(echo = F, warning = F, message = F, fig.align = "center",
                      fig.path = "assets/index_files/")
options(digits = 3)


## toolbox etc
library(ggplot2)
library(plotly)


## read data
base    <- readRDS("./data/base.rds")


## prepare cosmetics
source("./assets/theme_tut.R")
theme_set(theme_tut(size = 12))
```

# Introduction

This is my attempt to keep track of my glucose levels, insulin & carbohydrate intakes and all that. The data usually gets updated within the first few days of a month.

My pump model is: **Medtronic MiniMed 640G**  
My glucose meter is: **Medtronic Contour next link**

Medtronic provides a pretty neat web interface to read and upload all data tracked by the pump. Even better though is the possibility to export an already very tidy _csv_ file (apart from a few minor quirks regarding the data itself) which makes importing to **R** pretty comfortable.


# Bread Exchange & Basal Rate

## BE-Factor

  Time Range  | IU/BE
--------------|--------
 0:00 -  8:00 | 1.5
 8:00 - 12:00 | 1.25
12:00 - 14:00 | 1.7
14:00 - 15:30 | 1.5
15:30 - 18:00 | 1.15
18:00 - 24:00 | 1.5

04. March '17: _maybe lower 1.7 to 1.55 from 12:00 to 14:00_


## Current Basal Rate Profile

```{r rahmen_base1, eval = F}
ggplot(base, aes(x = Uhrzeit, y = Rate_2, group = 1)) +
  geom_line(linejoin = "round") +
  geom_point(size = 3, color = "#66cc00") +
  labs(x = "Uhrzeit", y = "Basalrate (IE/h)", 
       title = "Basalratenprofil 1", 
       subtitle = paste("Tagesgesamtmenge:", sum(base$Rate_2), "IE"))
```

```{r rahmen_base2}
hover <- ~paste("until<b>", Uhrzeit, "</b>o'clock:",
                "<br /><b>", Rate_2, "IU</b>")

plot_ly(base, x = ~Uhrzeit/(60*60), y = ~Rate_2) %>%
  add_lines(color = I("#66cc00"), hoverinfo = "text", 
            text = hover, showlegend = F) %>% 
  add_markers(color = I("black"), hoverinfo = "none", size = I(8), 
              showlegend = F) %>% 
  add_markers(color = I("#66cc00"), size = I(6), 
              hoverinfo = "text", text = hover) %>% 
  layout(title = "Basal rate configuration 1", 
         xaxis = list(title = "Time"),
         yaxis = list(title = "Basal rate (IU/h)"), 
         hovermode = "closest", autosize = T, showlegend = F) %>% 
  config(displayModeBar = F)
```


# To-Do:

- figure out why basal rate config plot is broken when knitted
- percentage of Bolus for food and for correction in "Overviews - Averages" table
- maybe switch theme to `sandstone`
